---
title: "Getting Started with Myrconn.PetroVisor.Client"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with Myrconn.PetroVisor.Client}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Introduction

The `Myrconn.PetroVisor.Client` package provides a comprehensive R wrapper for the PetroVisor REST API, enabling you to interact with PetroVisor services directly from R. This package structures API calls into intuitive services:

- **Data Services** (`sp$data`): Load and save time series, depth, static, and PVT data
- **Repository Services** (`sp$items`): Manage entities, signals, units, and hierarchies
- **Logging Services** (`sp$logs`): Access and manage log entries
- **File Services** (`sp$files`): Handle file operations
- **Tag Entry Services** (`sp$tag_entries`): Work with tag entries
- **Machine Learning Services** (`sp$ml`): Train and use ML models

## Installation

You can install the package from GitHub using the `remotes` package:

```{r install}
# Install from GitHub
remotes::install_github("Datagration/petrovisor-r-api")
```

## Loading the Package

```{r setup}
library(Myrconn.PetroVisor.Client)
```

# Authentication

Before you can use any PetroVisor services, you need to authenticate. The package supports three authentication methods.

## Method 1: API Key Authentication

This is the simplest method if you have an API key:

```{r auth-key}
# Create an authentication service instance
auth_service <- AuthenticationService$new()

# Get access token using API key
token_response <- auth_service$get_access_token(
  key = "your_api_key_here",
  discovery_url = "https://identity.us1.petrovisor.com"
)

# Extract the token
access_token <- token_response$access_token
```

## Method 2: Username and Password

If you have username and password credentials:

```{r auth-password}
auth_service <- AuthenticationService$new()

token_response <- auth_service$get_access_token(
  username = "your_username",
  password = "your_password",
  discovery_url = "https://identity.us1.petrovisor.com"
)

access_token <- token_response$access_token
```

## Method 3: Refresh Token

If you have a refresh token from a previous session:

```{r auth-refresh}
auth_service <- AuthenticationService$new()

token_response <- auth_service$get_access_token(
  refresh_token = "your_refresh_token",
  discovery_url = "https://identity.us1.petrovisor.com"
)

access_token <- token_response$access_token
```

For more detailed information about authentication, see the [Authentication Guide](authentication.html) vignette.

# Creating a Service Provider

The `ServiceProvider` class is your main entry point for interacting with PetroVisor. It provides access to all available services and handles authentication automatically.

## Using Username and Password

The most common way to create a service provider:

```{r sp-credentials}
sp <- ServiceProvider$new(
  url = "https://identity.us1.petrovisor.com",
  workspace = "YourWorkspaceName",
  user = "your_username",
  password = "your_password"
)
```

## Using a Token

If you already have an access token:

```{r sp-token}
sp <- ServiceProvider$new(
  url = "https://identity.us1.petrovisor.com",
  workspace = "YourWorkspaceName",
  client_token = access_token
)
```

# Quick Tour of Services

Once you have a `ServiceProvider` instance, you can access all the services through its fields:

## Repository Service (`sp$items`)

Access and manage PetroVisor items like entities and signals:

```{r items-examples}
# Load all entity names
entity_names <- sp$items$load_names("Entity")

# Load all signal names
signal_names <- sp$items$load_names("Signal")

# Load a specific entity by name
entity <- sp$items$load("Entity", "WellName")

# Create a new entity
new_entity <- Entity$new(
  name = "TestWell01",
  entity_type_name = "Well",
  alias = "Test Well 01",
  is_opportunity = FALSE
)
sp$items$save("Entity", new_entity)

# Delete an entity
sp$items$delete("Entity", "TestWell01")
```

## Data Service (`sp$data`)

Load and save data from PetroVisor:

```{r data-examples}
# Parse signal strings (signal name with unit in brackets)
oil_rate <- sp$parse_signal("Oil Rate [bbl/d]")
gas_rate <- sp$parse_signal("Gas Rate [Mscf/d]")

# Load time series data
data <- sp$data$load_signals(
  entities = c("Well1", "Well2"),
  signals = list(oil_rate, gas_rate),
  time_increment = "Daily",
  time_start = "2024-01-01T00:00:00",
  time_end = "2024-12-31T23:59:59",
  reshape = TRUE
)

# The result contains TimeNumericData with columns:
# - scenario, date, entity, and one column per signal
head(data$TimeNumericData)
```

## Logging Service (`sp$logs`)

Access log entries:

```{r logs-examples}
# Create log entries
log_entry1 <- LogEntry$new(
  severity = "Debug",
  message = "Test log message",
  category = "R Test",
  item_type = "Unknown",
  item_change = "Other"
)

log_entry2 <- LogEntry$new(
  severity = "Information",
  message = "Another test message",
  category = "R Test",
  item_type = "Unknown",
  item_change = "Other"
)

# Save log entries (takes a list or vector)
result <- sp$logs$save(c(log_entry1, log_entry2))

# Get available log categories
categories <- sp$logs$load_categories()

# Load log entries with optional filters
all_logs <- sp$logs$load()

# Load with specific category filter
r_test_logs <- sp$logs$load(categories = c("R Test"))

# Load with severity filter
warnings <- sp$logs$load(severities = c("Warning"))

# Load latest 10 entries
recent_logs <- sp$logs$load(last_entries = 10)
```

## Utility Functions

The `ServiceProvider` also includes helpful utility functions:

### Unit Conversion

Convert values between units:

```{r unit-conversion}
# Convert a single value
converted <- sp$convert_unit(
  x = 100,
  source_unit = "bbl/d",
  target_unit = "m3/d"
)

# Convert a vector of values
values <- c(100, 200, 300)
converted_values <- sp$convert_unit(
  x = values,
  source_unit = "bbl/d",
  target_unit = "m3/d"
)

# Handle special units
percent_to_fraction <- sp$convert_unit(10, "%", " ")  # 0.1
fraction_to_percent <- sp$convert_unit(0.1, " ", "%")  # 10
```

### Signal Parsing

Parse signal strings that include units:

```{r signal-parsing}
# Parse a mapped signal string from PetroVisor
signal <- sp$parse_signal("Oil Production Rate [bbl/d]")

# Access components
signal$Signal  # "Oil Production Rate"
signal$Unit    # "bbl/d"

# Use in data operations
data <- sp$data$load_signals(
  entities = c("Well1"),
  signals = list(signal),
  time_increment = "Daily",
  time_start = "2024-01-01T00:00:00",
  time_end = "2024-01-31T23:59:59",
  reshape = TRUE
)
```

### Sending Email

Send emails using PetroVisor's email configuration:

```{r email}
sp$send_mail(
  address = "recipient@example.com",
  subject = "Analysis Complete",
  body = "The production analysis has been completed successfully."
)
```

# Working with Entities

Entities represent wells, fields, reservoirs, and other assets in PetroVisor.

## Creating Entities

```{r create-entity}
# Create a new entity
well <- Entity$new(
  name = "Well-001",
  entity_type_name = "Well",
  alias = "Production Well 001",
  is_opportunity = FALSE
)

# Save to PetroVisor
result <- sp$items$save("Entity", well)
```

## Loading Entities

```{r load-entity}
# Load a specific entity
well <- sp$items$load("Entity", "Well-001")

# Get all entity names
all_entities <- sp$items$load_names("Entity")
```

## Updating Entities

```{r update-entity}
# Load entity
well <- sp$items$load("Entity", "Well-001")

# Modify properties
well$alias <- "Updated Alias"

# Save changes
sp$items$save("Entity", well)
```

## Deleting Entities

```{r delete-entity}
# Delete an entity
sp$items$delete("Entity", "Well-001")
```

# Working with Signals

Signals represent measured or calculated values in PetroVisor. Creating signals requires specifying all required properties.

## Creating Signals

```{r create-signal}
# Create a comprehensive signal definition
oil_rate_signal <- Signal$new(
  name = "Oil Rate",
  short_name = "qo",
  measurement_name = "VolumetricFlowRate",
  storage_unit_name = "bbl/d",
  aggregation_type = "Sum",
  container_aggregation_type = "Sum",
  signal_type = "TimeDependent",
  default_color = 0,
  default_line_type = "Solid",
  setting_name = NULL,
  labels = list(),
  description = "Oil production rate"
)

# Save signal
sp$items$save("Signal", oil_rate_signal)
```

## Signal Types

Different signal types are used for different data:

- `"Static"`: Entity-level data that doesn't change
- `"TimeDependent"`: Numeric data that varies over time
- `"StringTimeDependent"`: Text data that varies over time
- `"DepthDependent"`: Numeric data that varies with depth
- `"StringDepthDependent"`: Text data that varies with depth
- `"PVT"`: Pressure-Volume-Temperature dependent data

## Loading Signals

```{r load-signal}
# Load a specific signal
signal <- sp$items$load("Signal", "Oil Rate")

# Get all signal names
all_signals <- sp$items$load_names("Signal")
```

# Working with Data

## Loading Different Data Types

### Static Data

```{r load-static}
# Load static (non-time-varying) data
static_data <- sp$data$load_signals(
  entities = c("Well-001", "Well-002"),
  signals = lapply(
    c("Latitude [deg]", "Longitude [deg]"),
    function(x) {
      sp$parse_signal(x)
    }
  ),
  reshape = TRUE
)

# Access the data
static_data$StaticNumericData
```

### Time Series Data

```{r load-time}
# Load daily time series data
time_data <- sp$data$load_signals(
  entities = c("Well-001"),
  signals = lapply(
    c("Oil Rate [bbl/d]", "Gas Rate [Mscf/d]"),
    function(x) {
      sp$parse_signal(x)
    }
  ),
  time_increment = "Daily",
  time_start = "2024-01-01T00:00:00",
  time_end = "2024-01-31T23:59:59",
  reshape = TRUE
)

# Access the data
time_data$TimeNumericData
```

### Depth Data

```{r load-depth}
# Load depth-dependent data
depth_data <- sp$data$load_signals(
  entities = c("Well-001"),
  signals = lapply(
    c("Porosity [frac]", "Permeability [mD]"),
    function(x) {
      sp$parse_signal(x)
    }
  ),
  depth_increment = "Meter",
  depth_start = 5000,
  depth_end = 7000,
  reshape = TRUE
)

# Access the data
depth_data$DepthNumericData
```

### PVT Data

```{r load-pvt}
# Load PVT data
pvt_data <- sp$data$load_signals(
  entities = c("Reservoir-A"),
  signals = lapply(
    c("Oil FVF [rb/stb]", "Gas FVF [rcf/scf]"),
    function(x) {
      sp$parse_signal(x)
    }
  ),
  pressure_unit = "psi",
  temperature_unit = "degC",
  reshape = TRUE
)

# Access the data
pvt_data$PVTNumericData
```

## Saving Data

### Saving Time Series Data

```{r save-time}
# Prepare data frame with correct structure
time_data <- data.frame(
  scenario = c("", ""),
  date = c("2024-01-01T00:00:00", "2024-01-02T00:00:00"),
  entity = c("Well-001", "Well-001"),
  oil_rate = c(500, 520),
  gas_rate = c(1500, 1560)
)

# Column names must match signal names (without units)
colnames(time_data) <- c(
  "scenario", "date", "entity",
  "Oil Rate", "Gas Rate"
)

# Save data
result <- sp$data$save_signals(
  "TimeNumeric",  # Data type
  time_data,
  signals = lapply(
    c("Oil Rate [bbl/d]", "Gas Rate [Mscf/d]"),
    function(x) {
      sp$parse_signal(x)
    }
  )
)
```

### Saving Static Data

```{r save-static}
# Prepare static data
static_data <- data.frame(
  scenario = c(""),
  entity = c("Well-001"),
  latitude = c(45.123),
  longitude = c(-110.456)
)

colnames(static_data) <- c(
  "scenario", "entity",
  "Latitude", "Longitude"
)

# Save static data
result <- sp$data$save_signals(
  "StaticNumeric",
  static_data,
  signals = lapply(
    c("Latitude [deg]", "Longitude [deg]"),
    function(x) {
      sp$parse_signal(x)
    }
  )
)
```

## Deleting Data

```{r delete-data}
# Delete signal data for specific entities
result <- sp$data$delete_signals(
  entities = c("Well-001", "Well-002"),
  signals = c("Oil Rate", "Gas Rate"),
  time_start = "2024-01-01T00:00:00",
  time_end = "2024-01-31T23:59:59"
)
```

# Working with Reference Tables

Reference tables are structured tabular data with defined schemas.

## Creating a Reference Table

```{r create-ref-table}
# Define columns
key_column <- ReferenceTableColumn$new(
  name = "ID",
  column_type = "Numeric",
  unit_name = " "
)

name_column <- ReferenceTableColumn$new(
  name = "Name",
  column_type = "String",
  unit_name = " "
)

value_column <- ReferenceTableColumn$new(
  name = "Value",
  column_type = "Numeric",
  unit_name = "m3"
)

# Create reference table
ref_table <- ReferenceTable$new(
  name = "Parameters",
  description = "Parameter lookup table",
  key = key_column,
  values = list(name_column, value_column)
)

# Save reference table definition
sp$items$save("ReferenceTable", ref_table)
```

## Saving Reference Table Data

```{r save-ref-data}
# Prepare data
ref_data <- data.frame(
  Entity = c(NA, "Well-001"),
  Timestamp = c(NA, "2024-01-01T00:00:00"),
  ID = c(1, 2),
  Name = c("Parameter1", "Parameter2"),
  Value = c(100, 200)
)

# Save data
result <- sp$data$save_reference_table("Parameters", ref_data)
```

## Loading Reference Table Data

```{r load-ref-data}
# Load reference table data
data <- sp$data$load_reference_table("Parameters")
```

# Common Patterns

## Handling Missing Values

Use `NA` or `NaN` for missing numeric values:

```{r missing-values}
# Data with missing values
data <- data.frame(
  scenario = c("", "", ""),
  date = c("2024-01-01T00:00:00", "2024-01-02T00:00:00", "2024-01-03T00:00:00"),
  entity = c("Well-001", "Well-001", "Well-001"),
  oil_rate = c(500, NA, 520),  # Missing value on Jan 2
  gas_rate = c(1500, 1550, NaN)  # NaN on Jan 3
)

colnames(data) <- c("scenario", "date", "entity", "Oil Rate", "Gas Rate")

# PetroVisor handles NA/NaN appropriately
result <- sp$data$save_signals("TimeNumeric", data, signals)
```

## Working with Multiple Entities

```{r multiple-entities}
# Get all entity names
all_entities <- sp$items$load_names("Entity")

# Filter to specific entity types (e.g., wells starting with "PROD-")
production_wells <- all_entities[grepl("^PROD-", all_entities)]

# Load data for all production wells
production_data <- sp$data$load_signals(
  entities = production_wells,
  signals = list(sp$parse_signal("Oil Rate [bbl/d]")),
  time_increment = "Monthly",
  time_start = "2024-01-01T00:00:00",
  time_end = "2024-12-31T23:59:59",
  reshape = TRUE
)
```

## Error Handling

Always wrap API calls in error handling:

```{r error-handling}
# Wrap operations in tryCatch
result <- tryCatch({
  sp$data$save_signals("TimeNumeric", my_data, my_signals)
}, error = function(e) {
  cat("Error saving data:", e$message, "\n")
  return(NULL)
})

if (!is.null(result) && result$status_code == 200) {
  cat("Data saved successfully\n")
}
```

# Next Steps

Now that you understand the basics, explore these guides for more detailed information:

- **[Authentication Guide](authentication.html)**: Deep dive into authentication methods and best practices
- **[Working with Data](working-with-data.html)**: Comprehensive guide to loading and saving different data types
- **[Repository Service](repository-service.html)**: Managing entities, signals, hierarchies, and more
- **[Machine Learning](machine-learning.html)**: Training models and making predictions

# Getting Help

- Check the function documentation: `?ServiceProvider`, `?DataServices`, etc.
- Browse the [reference documentation](../reference/index.html)
- Visit the [GitHub repository](https://github.com/Datagration/petrovisor-r-api) to report issues or request features

# Important Notes

## Date/Time Format

Always use ISO 8601 format for dates and times:
- Correct: `"2024-01-01T00:00:00"`
- Incorrect: `"2024-01-01"` or `"01/01/2024"`

## Column Names in Data Frames

When saving data, column names in your data frame must match the signal names (without units):
- Signal: `"Oil Rate [bbl/d]"`
- Column name: `"Oil Rate"`

## Data Type Names

Use the correct data type names when saving:
- `"StaticNumeric"`, `"StaticString"`
- `"TimeNumeric"`, `"TimeString"`
- `"DepthNumeric"`, `"DepthString"`
- `"PVTNumeric"`

## Units with Special Characters

Some units require special handling:
- Space (dimensionless): `" "`
- Percent: `"%"`
- Units with slashes: handled automatically (e.g., `"bbl/d"`)
