---
title: "Machine Learning with PetroVisor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Machine Learning with PetroVisor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(Myrconn.PetroVisor.Client)
```

# Overview

The `MlTrainingService` provides machine learning capabilities within PetroVisor. You can train models, make predictions, and manage ML workflows directly from R. This guide covers the complete ML workflow with tested examples.

# Setup

Create a service provider to access the ML service:

```{r service-provider}
sp <- ServiceProvider$new(
  url = "https://identity.us1.petrovisor.com",
  workspace = "YourWorkspace",
  user = "your_username",
  password = "your_password"
)

# The ML service is accessible via sp$ml
```

# Machine Learning Workflow

The typical ML workflow in PetroVisor:

1. **Prepare Data**: Create entities and save signal data
2. **Create Model**: Define model with table formula and label
3. **Configure Training**: Set training options with scope and entity set
4. **Train**: Submit training request
5. **Monitor**: Check training status
6. **Evaluate**: Review training results
7. **Save**: Save the best trained model
8. **Predict**: Use the model for predictions

# Data Preparation

Before training, you need entities and signals with data.

## Creating Test Entities and Signals

```{r prepare-data}
# Create entities for training
entity_names <- c()
for (i in 1:50) {
  entity_name <- paste0("MLTestEntity_", sprintf("%02d", i))
  entity_names <- c(entity_names, entity_name)

  entity <- Entity$new(
    name = entity_name,
    entity_type_name = "Well",
    alias = paste0("MLE", sprintf("%02d", i)),
    is_opportunity = FALSE
  )

  sp$items$save("Entity", entity)
}

# Create predictor signal (input feature)
predictor_signal <- Signal$new(
  name = "ml test predictor signal",
  short_name = "mlpred",
  measurement_name = "Length",
  storage_unit_name = "m",
  aggregation_type = "Sum",
  container_aggregation_type = "Sum",
  signal_type = "Static",
  default_color = 3711337,
  default_line_type = "Solid",
  setting_name = NULL,
  labels = list(),
  description = "ML test predictor feature"
)

sp$items$save("Signal", predictor_signal)

# Create target signal (output to predict)
target_signal <- Signal$new(
  name = "ml test target signal",
  short_name = "mltarget",
  measurement_name = "Volume",
  storage_unit_name = "m3",
  aggregation_type = "Sum",
  container_aggregation_type = "Sum",
  signal_type = "Static",
  default_color = 16711680,
  default_line_type = "Solid",
  setting_name = NULL,
  labels = list(),
  description = "ML test target to predict"
)

sp$items$save("Signal", target_signal)

# Generate correlated data
set.seed(42)
predictor_values <- runif(50, min = 10, max = 100)
noise <- rnorm(50, mean = 0, sd = 5)
target_values <- 2.5 * predictor_values + 50 + noise

# Save data
static_data <- data.frame(
  scenario = rep("", 50),
  entity = entity_names,
  stringsAsFactors = FALSE
)
static_data[[predictor_signal$name]] <- predictor_values
static_data[[target_signal$name]] <- target_values

result <- sp$data$save_signals(
  "StaticNumeric",
  static_data,
  signals = lapply(
    c(
      paste0(predictor_signal$name,
             " [",
             predictor_signal$storage_unit_name,
             "]"),
      paste0(target_signal$name,
             " [",
             target_signal$storage_unit_name,
             "]")
    ),
    function(x) {
      sp$parse_signal(x)
    }
  )
)
```

# Creating and Configuring Models

## Basic Model Creation

```{r basic-model}
# Create table formula - defines features and target
table_formula <- paste0('Table "ML Training Data"
  Column "predictor" in "', predictor_signal$storage_unit_name, '"
    "', predictor_signal$name, '" in "', predictor_signal$storage_unit_name, '"
  End Column
  Column "target" in "', target_signal$storage_unit_name, '"
    "', target_signal$name, '" in "', target_signal$storage_unit_name, '"
  End Column
End Table')

# Create model
model <- MLModel$new(
  name = "Oil Rate Predictor",
  type = "Regression",
  table_formula = table_formula,
  label_column_name = "target"
)

# Model is not ready to train yet - need to set training options
```

## Creating Scope and Entity Set

Before setting training options, create scope and entity set:

```{r create-scope-entity-set}
# Create entity set
entity_set <- EntitySet$new(
  name = "ML Training Entities",
  entities = list(),  # Can provide entity objects or leave empty
  formula = paste0('Entity Set "ML Training Entities"
\t"MLTestEntity_01"
\t"MLTestEntity_02"
\t"MLTestEntity_03"
End Set')
)

# Create scope
scope <- Scope$new(
  name = "ML Training Scope",
  start = "2024-01-01T00:00:00.000Z",
  end = "2024-01-01T00:00:00.000Z",
  time_increment = "Daily",
  depth_increment = "Meter",
  start_depth = 0,
  end_depth = 5000,
  formula = 'Scope "At Date"
\tBetween #01/01/2024#
\tAnd #01/01/2024#
\tStep Daily
End Scope'
)
```

## Setting Training Options

Training options must be set before other configurations:

```{r set-training-options}
# Set training options (required before other model configurations)
model$set_training_options(
  scope = scope,
  entity_set = entity_set,
  test_fraction = 0.2,           # 20% for testing
  validation_fraction = 0.2,      # 20% for validation
  time_to_train = 60              # Training time in seconds
)
```

## Configuring Preprocessing

After setting training options, you can configure preprocessing:

```{r preprocessing}
# Enable specific preprocessing methods
model$set_preprocessing(
  c("MinMax", "BoxCox", "RobustScaling"),
  apply_pre_processors_before_training = FALSE
)
```

## Configuring Outlier Filtering

```{r outlier-filtering}
# Set outlier filtering method
model$set_outlier_filtering("CooksDistance")

# Options: "CooksDistance", "IQR", etc.
```

## Advanced: Neural Network Options

```{r neural-network}
# Configure neural network (after setting training options)
model$set_neural_network_options(
  library = "TensorFlow",
  optimizer = "Adam",
  learning_rate = 0.001,
  epochs = 100,
  batch_size = 64,
  dense_layers = 3,
  neurons = 128
)
```

## Advanced: Custom Training Options

```{r custom-training}
# Set custom training (e.g., GAM)
model$set_custom_training(
  training_type = "Gam",
  enable_pruning = TRUE,
  learning_rate = 0.05,
  maximum_bin_count_per_feature = 200,
  minimum_example_count_per_leaf = 10,
  max_iterations = 1500
)
```

## Saving Model Configuration

```{r save-model-config}
# Save the model configuration
result <- sp$items$save("MLModel", model)

if (result$status_code == 201) {
  cat("Model configuration saved successfully\n")
}
```

# Training Models

## Starting Training

```{r start-training}
# Train the model
training_request <- sp$ml$train(model)

# Get request ID for monitoring
request_id <- training_request$id
cat("Training started with request ID:", request_id, "\n")
```

## Monitoring Training Progress

```{r monitor-training}
# Check training status
status <- sp$ml$get_training_status(request_id)
cat("Status:", status$status, "\n")

# Status values: "Unknown", "Waiting", "Training", "Starting",
#                "Ready", "Completed", "Failed", "Canceled", "Canceling"

# Poll for completion
wait_for_training <- function(sp, request_id, max_wait = 600, interval = 30) {
  start_time <- Sys.time()

  repeat {
    status <- sp$ml$get_training_status(request_id)
    cat("Status:", status$status, "\n")

    if (status$status == "Ready") {
      cat("Training completed successfully!\n")
      break
    }

    if (status$status %in% c("Failed", "Canceled")) {
      stop("Training failed or was canceled")
    }

    elapsed <- as.numeric(difftime(Sys.time(), start_time, units = "secs"))
    if (elapsed > max_wait) {
      stop("Training timed out")
    }

    Sys.sleep(interval)
  }
}

# Wait for training to complete
wait_for_training(sp, request_id)
```

## Getting Training Results

```{r training-results}
# Get detailed training results (after training completes)
results <- sp$ml$get_training_results(request_id)

# results is an MLTrainingResults object
cat("Training results retrieved\n")
```

## Evaluating Results

```{r evaluate-results}
# Get best performing models
best_models <- results$get_best_models()

# Check if we have any best models
if (length(best_models) > 0) {
  best <- best_models[[1]]

  cat("Best Model:\n")
  cat("  Trainer:", best$trainer_name, "\n")

  # Metrics are in the outcome$values list
  if (!is.null(best$outcome) && length(best$outcome$values) > 0) {
    cat("  Validation Metrics:\n")
    for (metric in best$outcome$values) {
      cat("    ", metric$name, ":", metric$value, "\n")
    }
  }

  # Test metrics are in outcome$test_values
  if (!is.null(best$outcome) && length(best$outcome$test_values) > 0) {
    cat("  Test Metrics:\n")
    for (metric in best$outcome$test_values) {
      cat("    ", metric$name, ":", metric$value, "\n")
    }
  }
}
```

## Saving the Best Model

```{r save-best-model}
# Save the best model to PetroVisor
save_result <- sp$ml$save_best_model(
  model = model,
  best_models = best_models,
  request_id = request_id,
  rank = 1  # Save the top-ranked model
)

cat("Best model saved\n")
```

# Making Predictions

## Preparing Prediction Data

```{r prep-prediction}
# Prepare input values for prediction
# Must match the features used in training
prediction_values <- c(15, 25, 35, 45, 55)
```

## Making Predictions

```{r make-predictions}
# Make predictions using trained model
predictions <- sp$ml$predict(
  model_name = model$name,
  data = prediction_values
)

# predictions is a list where each element has:
# - Value: predicted value
# - Score: confidence score
# - Probability: probability (for classification)

# Extract predicted values
predicted_values <- sapply(predictions, function(p) p$Value)
cat("Predictions:", predicted_values, "\n")
```

## Batch Predictions

```{r batch-predictions}
# Load historical data
historical_data <- sp$data$load_signals(
  entities = c("MLTestEntity_01", "MLTestEntity_02"),
  signals = list(sp$parse_signal(paste0(predictor_signal$name, " [m]"))),
  reshape = TRUE
)

# Extract predictor values
predictor_vals <- historical_data$StaticNumericData[[predictor_signal$name]]

# Make predictions
batch_predictions <- sp$ml$predict(
  model_name = model$name,
  data = predictor_vals
)

# Extract values
predicted <- sapply(batch_predictions, function(p) p$Value)
```

# Model Management

## Loading Trained Models

```{r load-model}
# Load a saved model from repository
trained_model <- sp$items$load("MLModel", "Oil Rate Predictor")

# Check if model is trained
if (!is.null(trained_model$trained_models) &&
      length(trained_model$trained_models) > 0) {
  cat("Model is trained and ready for predictions\n")
}
```

## Listing Models

```{r list-models}
# Get all ML model names
model_names <- sp$items$load_names("MLModel")

# Load and summarize models
for (name in model_names) {
  model <- sp$items$load("MLModel", name)
  cat("Model:", model$name, "\n")
  cat("  Type:", model$type, "\n")
  cat("  Trained:", !is.null(model$trained_models) &&
        length(model$trained_models) > 0, "\n\n")
}
```

## Deleting Models

```{r delete-model}
# Delete a model
sp$items$delete("MLModel", "Old Model Name")
```

# Model Types

PetroVisor supports different model types:

```{r model-types}
# Regression (default)
regression_model <- MLModel$new(
  name = "Regression Model",
  type = "Regression",
  table_formula = table_formula,
  label_column_name = "target"
)

# Binary Classification
binary_model <- MLModel$new(
  name = "Binary Classification Model",
  type = "BinaryClassification",
  table_formula = table_formula,
  label_column_name = "category"
)

# Multiple Classification
multi_model <- MLModel$new(
  name = "Multi-Class Model",
  type = "MultipleClassification",
  table_formula = table_formula,
  label_column_name = "class"
)

# Clustering (no label needed)
cluster_model <- MLModel$new(
  name = "Clustering Model",
  type = "Clustering",
  table_formula = table_formula
)
```

# Complete Example

Here's a complete end-to-end example:

```{r complete-example}
# 1. Create and save model configuration
model <- MLModel$new(
  name = "Complete Example Model",
  type = "Regression",
  table_formula = table_formula,
  label_column_name = "target"
)

# 2. Configure model
model$set_training_options(scope, entity_set,
                           test_fraction = 0.2,
                           validation_fraction = 0.2,
                           time_to_train = 60)

model$set_preprocessing(
  methods = c("MinMax", "BoxCox", "RobustScaling"),
  apply_pre_processors_before_training = FALSE
)
model$set_outlier_filtering("CooksDistance")

# 3. Save configuration
sp$items$save("MLModel", model)

# 4. Train model
request <- sp$ml$train(model)
cat("Training request ID:", request$id, "\n")

# 5. Wait for training
wait_for_training(sp, request$id)

# 6. Get results
results <- sp$ml$get_training_results(request$id)
best_models <- results$get_best_models()

# 7. Save best model
sp$ml$save_best_model(model, best_models, request$id)

# 8. Make predictions
test_values <- c(20, 40, 60, 80)
predictions <- sp$ml$predict(model$name, test_values)

# 9. Extract and use predictions
predicted_values <- sapply(predictions, function(p) p$Value)
cat("Predictions:", predicted_values, "\n")
```

# Best Practices

## 1. Use Sufficient Training Data

```{r best-data}
# Ensure you have enough data points
# Rule of thumb: at least 10x the number of features
# For 5 features, aim for 50+ training examples
```

## 2. Split Data Appropriately

```{r best-split}
# Use typical splits:
model$set_training_options(
  scope, entity_set,
  test_fraction = 0.2,        # 20% for testing
  validation_fraction = 0.2,  # 20% for validation
  time_to_train = 300         # Adequate training time
)
```

## 3. Handle Training Errors

```{r best-errors}
# Wrap training in error handling
training_result <- tryCatch({
  request <- sp$ml$train(model)
  wait_for_training(sp, request$id)
  sp$ml$get_training_results(request$id)
}, error = function(e) {
  cat("Training failed:", e$message, "\n")
  return(NULL)
})

if (!is.null(training_result)) {
  # Process results
  best_models <- training_result$get_best_models()
}
```

## 4. Version Your Models

```{r best-versioning}
# Use naming conventions for model versions
model_name <- paste0("OilRatePredictor_v", format(Sys.Date(), "%Y%m%d"))

model <- MLModel$new(
  name = model_name,
  type = "Regression",
  table_formula = table_formula,
  label_column_name = "target",
  description = paste("Trained on", Sys.Date())
)
```

## 5. Document Models

```{r best-documentation}
# Include comprehensive descriptions
model <- MLModel$new(
  name = "Production Predictor v1",
  type = "Regression",
  table_formula = table_formula,
  label_column_name = "target",
  description = paste(
    "Predicts oil production rate from pressure and temperature.",
    "Features: Pressure, Temperature.",
    "Training data: 2024, 50 wells.",
    "Created:", Sys.Date()
  )
)
```

# Troubleshooting

## Issue: "You must call set_training_options first"

**Cause**: Trying to set preprocessing/custom training before training options

**Solution**: Always call `set_training_options()` before other configurations

```{r fix-order}
# Correct order:
model$set_training_options(scope, entity_set)  # First
model$set_preprocessing(...)                    # Then other configs
```

## Issue: Training fails with insufficient data

**Cause**: Not enough training examples

**Solution**: Ensure sufficient data points and check data quality

```{r fix-data}
# Check data availability
data <- sp$data$load_signals(entities, signals, reshape = TRUE)
cat("Data points:", nrow(data$StaticNumericData), "\n")
```

## Issue: Predictions are unrealistic

**Cause**: Input values outside training range

**Solution**: Validate prediction inputs

```{r fix-predictions}
# Check that prediction inputs are within training data range
training_range <- range(predictor_values, na.rm = TRUE)
prediction_range <- range(test_values, na.rm = TRUE)

if (prediction_range[1] < training_range[1] ||
      prediction_range[2] > training_range[2]) {
  warning("Prediction inputs outside training data range")
}
```

# Next Steps

- **[Working with Data](working-with-data.html)**: Learn about loading training data
- **[Repository Service](repository-service.html)**: Manage ML models and related items
- **[Getting Started](getting-started.html)**: Package overview

# See Also

- `?MlTrainingService`: ML training service documentation
- `?MLModel`: ML model class documentation
- Test files in `tests/testthat/test_machine_learning.R` and 
  `tests/testthat/test_ml_training_service.R` for working examples
