---
title: "Managing Items with Repository Service"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Managing Items with Repository Service}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(Myrconn.PetroVisor.Client)
```

# Overview

The `RepositoryService` provides access to PetroVisor's repository of items, including entities, signals, units, hierarchies, and more. This service allows you to list, load, create, update, and delete these items.

# Setup

Create a service provider to access the repository service:

```{r service-provider}
sp <- ServiceProvider$new(
  url = "https://identity.us1.petrovisor.com",
  workspace = "YourWorkspace",
  user = "your_username",
  password = "your_password"
)

# The repository service is accessible via sp$items
```

# Core Operations

The repository service provides four main operations for all item types:

1. **`load_names(type)`**: Get a list of all item names
2. **`load(type, name)`**: Retrieve a specific item by name
3. **`save(type, object)`**: Create or update an item
4. **`delete(type, name)`**: Remove an item by name

These are the **only** methods available - there are no filtering or search methods.

## Loading Item Names

Get all names for a specific item type:

```{r load-names}
# Load all entity names
entity_names <- sp$items$load_names("Entity")

# Load all signal names
signal_names <- sp$items$load_names("Signal")

# Load all unit names
unit_names <- sp$items$load_names("Unit")

# Load all hierarchy names
hierarchy_names <- sp$items$load_names("Hierarchy")
```

## Loading Specific Items

Load a specific item by name:

```{r load-item}
# Load a specific entity
well <- sp$items$load("Entity", "Well-001")

# Load a specific signal
oil_rate_signal <- sp$items$load("Signal", "Oil Rate")

# Load a specific unit
unit <- sp$items$load("Unit", "bbl/d")
```

## Saving Items

Create or update items:

```{r save-item}
# Create a new entity
new_well <- Entity$new(
  name = "Well-123",
  entity_type_name = "Well",
  alias = "Production Well 123",
  is_opportunity = FALSE
)

# Save the entity (creates or updates)
result <- sp$items$save("Entity", new_well)

# Check result
if (result$status_code == 201) {
  cat("Entity created successfully\n")
}
```

## Deleting Items

Remove items by name:

```{r delete-item}
# Delete an entity
result <- sp$items$delete("Entity", "Well-123")

# Check result
if (result$status_code == 200) {
  cat("Entity deleted successfully\n")
}

# Trying to load deleted item will error
# sp$items$load("Entity", "Well-123")  # This will throw an error
```

# Working with Entities

Entities are the primary objects in PetroVisor (wells, fields, reservoirs, etc.).

## Creating Entities

```{r create-entity}
# Create a new well entity
well <- Entity$new(
  name = "Well-456",
  entity_type_name = "Well",
  alias = "Injection Well 456",
  is_opportunity = FALSE
)

# Save to PetroVisor
result <- sp$items$save("Entity", well)
```

## Loading Entities

```{r load-entities}
# Get all entity names
all_entities <- sp$items$load_names("Entity")

# Load a specific entity
well <- sp$items$load("Entity", "Well-001")

# Access entity properties
cat("Name:", well$name, "\n")
cat("Type:", well$entity_type_name, "\n")
cat("Alias:", well$alias, "\n")
```

## Updating Entities

```{r update-entity}
# Load existing entity
well <- sp$items$load("Entity", "Well-456")

# Modify properties
well$alias <- "Updated Alias"

# Save changes (same save method)
result <- sp$items$save("Entity", well)
```

## Filtering Entities (Manual)

Since there's no `load_by_filter()` method, you need to filter manually:

```{r filter-entities}
# Get all entity names
all_entities <- sp$items$load_names("Entity")

# Filter by pattern (e.g., wells starting with "PROD-")
production_wells <- all_entities[grepl("^PROD-", all_entities)]

# Filter to specific names
specific_wells <-
  all_entities[all_entities %in% c("Well-001", "Well-002", "Well-003")]

# Load filtered entities
well_objects <- lapply(production_wells, function(name) {
  sp$items$load("Entity", name)
})
```

# Working with Entity Types

Entity types define categories of entities (Well, Field, Reservoir, etc.).

## Loading Entity Types

```{r entity-types}
# Get all entity type names
entity_type_names <- sp$items$load_names("EntityType")

# Load a specific entity type
well_type <- sp$items$load("EntityType", "Well")

# Access properties
cat("Name:", well_type$name, "\n")
```

## Creating Custom Entity Types

```{r create-entity-type}
# Create a new entity type
custom_type <- EntityType$new(
  name = "CustomAsset",
  image = "default"  # Or base64 encoded image
)

# Save entity type
result <- sp$items$save("EntityType", custom_type)
```

# Working with Signals

Signals represent measured or calculated values in PetroVisor. **Creating signals requires ALL fields.**

## Loading Signals

```{r load-signals}
# Get all signal names
all_signals <- sp$items$load_names("Signal")

# Load a specific signal
oil_rate <- sp$items$load("Signal", "Oil Rate")

# Access signal properties
cat("Name:", oil_rate$name, "\n")
cat("Storage Unit:", oil_rate$storage_unit_name, "\n")
cat("Signal Type:", oil_rate$signal_type, "\n")
```

## Creating Signals

Signals require **all** of these fields:

```{r create-signal}
# Create a comprehensive signal definition
new_signal <- Signal$new(
  name = "Custom Production Rate",
  short_name = "cpr",
  measurement_name = "VolumetricFlowRate",
  storage_unit_name = "bbl/d",
  aggregation_type = "Sum",
  container_aggregation_type = "Sum",
  signal_type = "TimeDependent",
  default_color = 0,
  default_line_type = "Solid",
  setting_name = NULL,
  labels = list(),
  description = "Custom calculated production rate"
)

# Save signal
result <- sp$items$save("Signal", new_signal)
```

## Signal Types

Different signal types for different data:

- `"Static"`: Entity-level data that doesn't change
- `"TimeDependent"`: Numeric data varying over time
- `"StringTimeDependent"`: Text data varying over time
- `"DepthDependent"`: Numeric data varying with depth
- `"StringDepthDependent"`: Text data varying with depth
- `"PVT"`: Pressure-Volume-Temperature dependent data

## Signal Properties Reference

- **name**: Signal identifier (required)
- **short_name**: Abbreviated name (required)
- **measurement_name**: Type of measurement (e.g., "Length", "Volume", "Dimensionless") (required)
- **storage_unit_name**: Unit for storage (e.g., "m", "m3", " ") (required)
- **aggregation_type**: How to aggregate (e.g., "Sum", "Average") (required)
- **container_aggregation_type**: Container aggregation (required)
- **signal_type**: Type of signal (see above) (required)
- **default_color**: Numeric color code (required)
- **default_line_type**: Line style (e.g., "Solid", "Dash") (required)
- **setting_name**: Optional setting reference (can be NULL)
- **labels**: List of labels (can be empty list)
- **description**: Description text (required)

# Working with Units

Units define measurement systems in PetroVisor.

## Loading Units

```{r load-units}
# Get all unit names
all_units <- sp$items$load_names("Unit")

# Load a specific unit
bbl_d <- sp$items$load("Unit", "bbl/d")

# Access unit properties
cat("Name:", bbl_d$name, "\n")
cat("Measurement:", bbl_d$unit_measurement_name, "\n")
```

## Unit Measurements

Units belong to unit measurements:

```{r unit-measurements}
# Load all unit measurement names
measurements <- sp$items$load_names("UnitMeasurement")

# Load a specific measurement
flow_rate <- sp$items$load("UnitMeasurement", "VolumetricFlowRate")

# See all units in this measurement
cat("Units:", paste(flow_rate$unit_names, collapse = ", "), "\n")
```

# Working with Reference Tables

Reference tables define structured data schemas.

## Creating Reference Tables

```{r create-ref-table}
# Define columns
key_column <- ReferenceTableColumn$new(
  name = "ID",
  column_type = "Numeric",
  unit_name = " "
)

name_column <- ReferenceTableColumn$new(
  name = "Name",
  column_type = "String",
  unit_name = " "
)

value_column <- ReferenceTableColumn$new(
  name = "Value",
  column_type = "Numeric",
  unit_name = "m3"
)

# Create reference table
ref_table <- ReferenceTable$new(
  name = "Parameters",
  description = "Parameter lookup table",
  key = key_column,
  values = list(name_column, value_column)
)

# Save reference table definition
result <- sp$items$save("ReferenceTable", ref_table)
```

## Loading Reference Tables

```{r load-ref-tables}
# Get all reference table names
ref_table_names <- sp$items$load_names("ReferenceTable")

# Load a specific reference table definition
ref_table <- sp$items$load("ReferenceTable", "Parameters")
```

# Working with Pivot Tables

Pivot tables define pre-aggregated summary tables.

## Creating Pivot Tables

```{r create-pivot}
# Create a pivot table definition
pivot_table <- PivotTable$new(
  name = "Monthly Production Summary",
  add_entity_alias_column = TRUE,
  add_entity_type_column = TRUE,
  scope_formula = 'Scope "Monthly"
\tBetween #01/01/2024#
\tAnd #12/31/2024#
\tStep Monthly
End Scope',
  entity_set_formula = 'Entity Set "Production Wells"
\t"Well-001"
\t"Well-002"
End Set',
  table_formula = 'Table "Production"
\tColumn "Oil" in "bbl/d"
\t\t"Oil Rate" in "bbl/d"
\tEnd Column
End Table',
  skip_empty_rows = TRUE,
  add_is_opportunity_column = FALSE,
  append_data = FALSE,
  description = "Monthly production summary"
)

# Save pivot table definition
result <- sp$items$save("PivotTable", pivot_table)
```

## Loading Pivot Tables

```{r load-pivot}
# Get all pivot table names
pivot_names <- sp$items$load_names("PivotTable")

# Load a specific pivot table definition
pivot <- sp$items$load("PivotTable", "Monthly Production Summary")
```

# Advanced Operations

## Batch Loading

Load multiple items efficiently:

```{r batch-load}
# Get all entity names
entity_names <- sp$items$load_names("Entity")

# Load first 10 entities
entities <- lapply(entity_names[1:10], function(name) {
  sp$items$load("Entity", name)
})

# Extract information
entity_info <- do.call(rbind, lapply(entities, function(e) {
  data.frame(
    name = e$name,
    type = e$entity_type_name,
    alias = if (is.null(e$alias)) NA else e$alias,
    stringsAsFactors = FALSE
  )
}))

print(entity_info)
```

## Searching and Filtering

Manual filtering using R:

```{r filtering}
# Load all entity names
all_entities <- sp$items$load_names("Entity")

# Filter by pattern
wells <- all_entities[grepl("^Well-", all_entities)]
fields <- all_entities[grepl("Field", all_entities)]

# Filter by multiple patterns
production_assets <- all_entities[grepl("PROD|Well", all_entities)]

# Case-insensitive search
search_term <- "test"
matching_entities <-
  all_entities[grepl(search_term, all_entities, ignore.case = TRUE)]
```

## Exporting Repository Structure

```{r export-structure}
# Function to export entity structure
export_entities <- function(sp) {
  # Get all entities
  entity_names <- sp$items$load_names("Entity")

  # Load all entities
  entities <- lapply(entity_names, function(name) {
    tryCatch({
      sp$items$load("Entity", name)
    }, error = function(e) {
      NULL
    })
  })

  # Remove NULLs (failed loads)
  entities <- entities[!sapply(entities, is.null)]

  # Convert to data frame
  entity_df <- do.call(rbind, lapply(entities, function(e) {
    data.frame(
      name = e$name,
      type = e$entity_type_name,
      alias = if (is.null(e$alias)) "" else e$alias,
      is_opportunity = e$is_opportunity,
      stringsAsFactors = FALSE
    )
  }))

  return(entity_df)
}

# Export to CSV
entity_export <- export_entities(sp)
write.csv(entity_export, "entities.csv", row.names = FALSE)
```

## Bulk Creation

```{r bulk-create}
# Create multiple entities from a data frame
entity_data <- data.frame(
  name = c("Well-201", "Well-202", "Well-203"),
  type = c("Well", "Well", "Well"),
  alias = c("Producer 201", "Producer 202", "Injector 203"),
  stringsAsFactors = FALSE
)

# Create and save entities
for (i in 1:nrow(entity_data)) {
  entity <- Entity$new(
    name = entity_data$name[i],
    entity_type_name = entity_data$type[i],
    alias = entity_data$alias[i],
    is_opportunity = FALSE
  )

  result <- sp$items$save("Entity", entity)

  if (result$status_code == 201) {
    cat("Created:", entity_data$name[i], "\n")
  }
}
```

# Best Practices

## 1. Check Before Deleting

```{r best-delete}
# Check if item exists before deleting
item_names <- sp$items$load_names("Entity")

if ("Well-123" %in% item_names) {
  sp$items$delete("Entity", "Well-123")
  cat("Deleted Well-123\n")
} else {
  cat("Well-123 not found\n")
}
```

## 2. Handle Errors Gracefully

```{r best-errors}
# Wrap operations in tryCatch
result <- tryCatch({
  entity <- sp$items$load("Entity", "NonExistent")
  entity
}, error = function(e) {
  cat("Error loading entity:", e$message, "\n")
  return(NULL)
})

if (!is.null(result)) {
  # Process entity
  cat("Loaded:", result$name, "\n")
}
```

## 3. Validate Before Saving

```{r best-validate}
# Validate entity before saving
create_validated_entity <- function(name, type, alias) {
  # Check name
  if (is.null(name) || name == "") {
    stop("Entity name cannot be empty")
  }

  # Check type exists
  types <- sp$items$load_names("EntityType")
  if (!(type %in% types)) {
    stop("Invalid entity type: ", type)
  }

  # Create entity
  entity <- Entity$new(
    name = name,
    entity_type_name = type,
    alias = alias,
    is_opportunity = FALSE
  )

  return(entity)
}

# Use validation
entity <- create_validated_entity("Well-999", "Well", "Test Well")
result <- sp$items$save("Entity", entity)
```

## 4. Use Consistent Naming

```{r best-naming}
# Follow naming conventions
# Good examples:
# - "Well-001", "Well-002" (consistent numbering)
# - "Field-North", "Field-South" (descriptive)

# Create with naming function
create_well_entity <- function(number, area) {
  name <- sprintf("Well-%s-%03d", area, number)
  alias <- sprintf("%s Area Well %d", area, number)

  Entity$new(
    name = name,
    entity_type_name = "Well",
    alias = alias,
    is_opportunity = FALSE
  )
}

# Use consistent naming
well <- create_well_entity(1, "North")
result <- sp$items$save("Entity", well)
```

# Item Type Reference

Common item types you can work with:

- **Entity**: Wells, fields, reservoirs, etc.
- **EntityType**: Categories of entities
- **Signal**: Measured or calculated values
- **Unit**: Measurement units
- **UnitMeasurement**: Unit categories
- **Hierarchy**: Tree structures
- **ReferenceTable**: Structured tables
- **PivotTable**: Pre-aggregated tables
- **Scenario**: Data versions
- **Scope**: Entity and signal selections
- **Tag**: Metadata tags
- **MLModel**: Machine learning models
- **RScript**: R scripts
- **Workflow**: Automated processes

# Important Notes

## No Search or Filter Methods

The repository service does **NOT** have these methods:
- ❌ `load_by_filter()`
- ❌ `load_by_name()` (use `load()` instead)
- ❌ `search()`
- ❌ `find()`

You must use `load_names()` to get all names, then filter in R.

## Signal Requirements

When creating signals, you **must** provide all 13 required fields. Missing any field will cause an error.

## Case Sensitivity

Item names are case-sensitive. "Well-001" and "well-001" are different items.

# Next Steps

- **[Working with Data](working-with-data.html)**: Load and save data using repository items
- **[Machine Learning](machine-learning.html)**: Use entities and signals in ML models
- **[Getting Started](getting-started.html)**: Package overview

# See Also

- `?RepositoryService`: Complete repository service documentation
- `?Entity`: Entity class documentation
- `?Signal`: Signal class documentation
- `?ServiceProvider`: Service provider documentation
- Test files in `tests/testthat/test_entity.R` and `tests/testthat/test_signal.R` for examples
